---
title: "Interactive & choropleth mapping"
description: |
  In this post I use publically-available data on California Oil Spills to show how to quickly and easily create an interactive map using tmap. I also use these data to create a choropleth map showing the number of inland oil spills per county.
author:
  - name: Julianna Renzi
date: 02-24-2021
output:
  distill::distill_article:
    self_contained: false
    code_folding: true
    toc: true
    toc_float: true
---


```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
require(here)
require(sf)
require(tmap)
```


## Load the data

These data are on oil spills in California. They were downloaded as a shapefile from the **California Department of Fish and Wildlife** (read more about the dataset here: https://map.dfg.ca.gov/metadata/ds0394.html)

```{r, message=FALSE, warning=FALSE}
oil_spills <- read_sf(here("_posts", "2021-02-24-interactivemap", "data", "ds394", "ds394.shp"))

oil_spills %>% st_crs # NAD 1983
```

Read in county borders

```{r, message=FALSE, warning=FALSE}
ca_counties <- read_sf(here("_posts", "2021-02-24-interactivemap", "data", "ca_counties", "CA_Counties_TIGER2016.shp")) %>% 
  select(NAME, ALAND) %>% 
  rename(county_name = NAME, land_area = ALAND) %>% 
  st_transform(crs = st_crs(oil_spills)) # want crs to match oil spill data

# double check CRS
st_crs(oil_spills) == st_crs(ca_counties) # looks good
```


```{r, message=FALSE, warning=FALSE}
ggplot() +
  geom_sf(data = ca_counties) +
  geom_sf(data = oil_spills, aes(color = INLANDMARI), alpha = 0.5) +
  theme_minimal()
```

## Make an interactive map

Now make it interactive

```{r, message=FALSE, warning=FALSE}
# set viewing mode to interactive
tmap_mode(mode = "view")

# make map with the polygon fill color updated by variable 'land_area', updating the color palette to "BuGn"), then add another shape layer for the sesbania records (added as dots)
tm_shape(ca_counties) +
  tm_polygons(col = "white") +
  tm_shape(oil_spills) +
  tm_dots(col = "INLANDMARI", title = "Area") +
  tm_basemap("Esri.WorldTopoMap") # this sets the basemap--I changed it to the topographic map for fun but there are 3 options and one can change it in the GUI
```

## Create the choropleth map

Now make a choropleth map where the fill color for each county depends on the count of inland oil spill events for 2008

```{r, message=FALSE, warning=FALSE}
# join data points to counties
oil_counties <- ca_counties %>% 
  st_join(oil_spills)

oil_count <- oil_counties %>% 
  filter(INLANDMARI == "Inland") %>% 
  count(LOCALECOUN) 


# plot
ggplot(data = oil_count) +
  geom_sf(aes(fill = n), color = "white", size = 0.1) + 
  scale_fill_gradientn(colors = c("lightgray", "orange", "red")) +
  theme_minimal() +
  labs(fill = "Number of inland oil spills")
```


