---
title: CV
name: Julianna
surname: Renzi
position: "PhD Student"
address: "Marine Sciences Institute: University of California, Santa Barbara"
www: juliannarenzi.webflow.io/
email: "jrenzi[at]ucsb.edu"
twitter: JuliannaRenzi
github: juliannajollyrenzi
linkedin: juliannarenzi
date: "`r format(Sys.time(), '%B %Y')`"
aboutme: "Julianna is a first year PhD student at the University of California, Santa Barbara."
output: 
  vitae::markdowncv:
    theme: kjhealy
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(vitae)
```

## About

I study community ecology and the role of biodiversity in ecosystem stability. I work in coastal environments, with a current focus on coral reefs, although I've conducted research in salt marsh, oyster reef, mangrove, and desert ecosystems. I'm currently a first year PhD student in the Burkepile lab at UC Santa Barbara where I am studying the coral reefs of Mo'orea, French Polynesia. I love coding, data analysis, and Thai food. 

## Education

```{r}
library(tibble)
tribble(
  ~ Degree, ~ Year, ~ Institution, ~ Where,
  "Informal studies", "1889-91", "Flying University", "Warsaw, Poland",
  "Master of Physics", "1893", "Sorbonne Université", "Paris, France",
  "Master of Mathematics", "1894", "Sorbonne Université", "Paris, France"
) %>% 
  detailed_entries(Degree, Year, Institution, Where)
```

## Nobel Prizes

```{r}
tribble(
  ~Year, ~Type, ~Desc,
  1903, "Physics", "Awarded for her work on radioactivity with Pierre Curie and Henri Becquerel",
  1911, "Chemistry", "Awarded for the discovery of radium and polonium"
) %>% 
  brief_entries(
    glue::glue("Nobel Prize in {Type}"),
    Year, 
    Desc
  )
```

## Publications

```{r}
library(dplyr)
knitr::write_bib(c("vitae", "tibble"), "packages.bib")

bibliography_entries("packages.bib") %>%
  arrange(desc(author$family), issued)
```

