---
title: "Working with rasters"
description: |
  In this post I use the output of a model describing the likelihood of Californian cetacean presence to create a raster of predicted species diversity off the coast of southern California.
author:
  - name: Julianna Renzi
    url: {}
date: 02-24-2021
output:
  distill::distill_article:
    self_contained: false
    code_folding: true
    toc: true
    toc_float: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(raster)
library(sf)
library(fasterize) # use sf geometries and rasterize them efficiently
library(rgdal)
library(rnaturalearth) # get coastline of California
library(rnaturalearthdata)
library(rgeos)
library(ggsn)
```

## Read in the data

Read in some raster data on California cetaceans, which I'll use in this example. These data are taken from **AquaMaps** (*Kaschner, K., Rius-Barile, J., Kesner-Reyes, K., Garilao, C., Kullander, S., Rees, T., & Froese, R. (2016). AquaMaps: Predicted range maps for aquatic species. www.aquamaps.org*). The data cover 35 Californian cetacean species and the probability of seeing each species at a given location given environmental factors such as depth, temperature, salinity, and proximity to shore. To see the code I used in each section click on the `Show code` option.

```{r}
# get the names of rasters for each species
cet_tifs <- list.files(path = here("_posts", "2021-02-24-workingwithrasters", "ca_cetaceans"), pattern = ".tif", full.names = TRUE)

# bring rasters in as a raster stack
cet_stack <- raster::stack(cet_tifs)

# can view the object to make sure it looks correct:
cet_stack # looks good

```

Plot the predicted distribution of a few species to examine the data. First for *Balaenoptera acutorostrata*:

```{r}
plot(cet_stack$Balaenoptera_acutorostrata, col = hcl.colors(n = 100))
title(expression(paste("Probability of occurence of ", italic("Balaenoptera acutorostrata"))), cex.main = 0.8,
      xlab = "Latitude", ylab = "Longitude")
```

Then for *Globicephala macrorhynchus*:

```{r}
plot(cet_stack$Globicephala_macrorhynchus, col = hcl.colors(n = 100))
title(expression(paste("Probability of occurence of ", italic("Globicephala macrorhynchus"))), cex.main = 0.8,
      xlab = "Latitude", ylab = "Longitude")
```

Then for *Kogia sima*

```{r}
plot(cet_stack$Kogia_sima, col = hcl.colors(n = 100))
title(expression(paste("Probability of occurence of ", italic("Kogia sima"))), cex.main = 0.8,
      xlab = "Latitude", ylab = "Longitude")
```


## Turn the data into presence/absence using raster algebra

For the purposes of this exercise, we'll define a species as likely to be "present" when the probability of observing it is greater than 0.75. 

```{r}
convert_pa <- function(x) {
  ifelse(x > 0.75, 1, 0)
}

cet_presence <- calc(cet_stack, function(x){convert_pa(x)})

```

Then we want to compare these plots with the modeled data above to make sure our function worked the way we wanted it to.

Here's the output for *Balaenoptera acutorostrata*:

```{r}
plot(cet_presence$Balaenoptera_acutorostrata, col = hcl.colors(n = 2))
title(expression(paste("Presence/absence of ", italic("Balaenoptera acutorostrata"))), cex.main = 0.8,
      xlab = "Latitude", ylab = "Longitude")
```

And for *Globicephala macrorhynchus*

```{r}
plot(cet_presence$Globicephala_macrorhynchus, col = hcl.colors(n = 2))
title(expression(paste("Presence/absence of ", italic("Globicephala macrorhynchus"))), cex.main = 0.8,
      xlab = "Latitude", ylab = "Longitude")
```

And for *Kogia sima*

```{r}
plot(cet_presence$Kogia_sima, col = hcl.colors(n = 2))
title(expression(paste("Presence/absence of ", italic("Kogia sima"))), cex.main = 0.8,
      xlab = "Latitude", ylab = "Longitude")
# all look reasonable given the raw maps above
```

All of which look reasonable given the raw maps above. Nice!

## Calculate and plot potential species richness

Here we sum the "presence" of each organism in each cell and plot those results as an approximation of species richness.

```{r}
cet_sum <- calc(cet_presence, sum, na.rm = FALSE) # get the sum of 1's (presence as we defined it) for the number of potential species in a grid cell

# can see what the summary looks like
# plot(cet_sum, col = hcl.colors(n = 20))

# get it into data.frame format for ggplot
cet_df <- raster::rasterToPoints(cet_sum) %>% # this is WGS84
  as.data.frame()

# get coastline data for reference
coasts <- ne_coastline(scale = "medium", returnclass = "sf")

# can visualize what the map looks like before using by running this code:
# plot(coasts) # this is WGS84

# plot with ggplot()
ggplot(data = coasts) +
  geom_sf() +
  coord_sf(xlim = c(-125, -113), ylim = c(28, 38), expand = FALSE) +
  annotate(geom = "text", x = -121, y = 30, label = "Pacific Ocean", 
     fontface = "italic", color = "grey22", size = 3) +
  annotate(geom = "text", x = -118, y = 36, label = "California", 
           color = "grey22", size = 3) +
  geom_raster(data = cet_df, aes(x = x, y = y, fill = layer), alpha = 0.75) +
  scale_fill_gradient(low = "lightblue", high = "darkred",
                           breaks=seq(from = 1, to = 25, by = 2)) +
  theme_bw() +
  theme(axis.title.x = element_blank(), 
     axis.title.y = element_blank(), 
     panel.border = element_rect(fill = NA)) +
  labs(fill = "Species richness",
       caption = expression(atop("Map of predicted cetacean species richness", "in the California Bight where a prediction probability of", "0.75 was considered a presence and below that was an absence")))
```

